The powder x-ray diffractogram open database

### 2024-04-29 paper/01_introduction.tex
Commited changes by Henrik

@@

Secondly we would like to also present this dataset as a challenge: The fraction of labeled data can be used to
easily numerically evalute the performance of the model on a broad range of data as it appears in experiment, which can be used
to gauge how well models would perform in practice. \\
But even the unlabeled portion of the dataset can be used to test the performance of models: The least a model
has to do in order to prove useful in practice is to make a guess about the crystal structure that is good enough
for Rietveld refinement to converge toward the true structure. \\
Hence, the performance of the model can be tested by evaluating if the inital guess of the model is sufficient for
Rietveld refinment to refine the initial guess to a structure that produces a diffractogram matching the one observed in experiment.
Finally, by inviting the application of models to data as it appears in experiment we would like to bring attention
to some of the practical difficulties that need to be addressed in applying a model to a wide variety of x-ray diffractograms
collected from different sources: For example the angle and the step size of the diffractograms will differ between sources
, which is something that any model that aims to perform well on this dataset must be equipped to handle.

>>>>

\begin{itemize}
    \item Plan for Figures / Tables
    \begin{itemize}
        \item Figure 1: Multiple panels showing histograms of some key metrics: Space group, unit cell volume (normalizable?), ...
        \item Table 1: List of existing powder XRD datasets
        \item Table 2: List of existing ML approaches for pXRD
        \item Table 3: New pXRD datasets offered by this paper
    \end{itemize}

    \item Additional thoughts
    \begin{itemize}
        \item General structure of the paper:
        \begin{itemize}
            \item Introduction
            \item Introduce existing datasets and already show why they are not enough / too small
            \item Summarize SOTA ML that has been applied to experimental pXRD, discuss the limits and gap between performance on simulated data and exp. data (find some good example papers where this is very apparent)
            \item Summary of our dataset, including paragraphs of the different contributions by other groups; comparison with existing datasets, especially in terms of size
        \end{itemize} 
    \end{itemize}
\end{itemize}

@@

Secondly we would like to also present this dataset as a challenge: The fraction of labeled data can be used to
easily numerically evalute the performance of the model on a broad range of data as it appears in experiment, which can be used
to gauge how well models would perform in practice. \\
But even the unlabeled portion of the dataset can be used to test the performance of models: The least a model
has to do in order to prove useful in practice is to make a guess about the crystal structure that is good enough
for Rietveld refinement to converge toward the true structure. \\
Hence, the performance of the model can be tested by evaluating if the inital guess of the model is sufficient for
Rietveld refinment to refine the initial guess to a structure that produces a diffractogram matching the one observed in experiment.
Finally, by inviting the application of models to data as it appears in experiment we would like to bring attention
to some of the practical difficulties that need to be addressed in applying a model to a wide variety of x-ray diffractograms
collected from different sources: For example the angle and the step size of the diffractograms will differ between sources, 
which is something that any model that aims to perform well on this dataset must be equipped to handle.

### 2024-04-29 paper/02_existing_datasets.tex
Commited changes by Henrik

% -> PDF (https://www.icdd.com/) -> 
%- 1200k "material data sets" (structures?) 
% + 19 k entries with "Experimental raw data digital patterns"; 
%  Proprietary
%The PDF itself lists the following data sources:
\begin{itemize}
    \item \textbf{ICDD Powders (00) – International Centre for Diffraction Data Newtown Square, PA, USA}
    \item \textbf{ICSD (01) – FIZ Karlsruhe – Leibniz Institute for Information Infrastructure (FIZ), Karlsruhe, Germany – Inorganic Crystal Structure Database}
    \item \textbf{NIST (03) – National Institute of Standards and Technology, Gaithersburg, Maryland, USA}
    \item \textbf{MPDS (04) – Material Phases Data System, Vitznau, Switzerland – Linus Pauling File (LPF)}
    \item \textbf{ICDD Single Crystal Data (05) – International Centre for Diffraction Data Newtown Square, PA, USA}
\end{itemize}

>>>>

% -> PDF (https://www.icdd.com/) -> 
%- 1200k "material data sets" (structures?) 
% + 19 k entries with "Experimental raw data digital patterns"; 
%  Proprietary
%The PDF itself lists the following data sources:

\begin{itemize}
\item Powder Diffraction File (ICDD)
\begin{itemize}
    \item ICDD Powders (00) – International Centre for Diffraction Data Newtown Square, PA, USA
    \item ICSD (01) – FIZ Karlsruhe – Leibniz Institute for Information Infrastructure (FIZ), Karlsruhe, Germany – Inorganic Crystal Structure Database
    \item NIST (03) – National Institute of Standards and Technology, Gaithersburg, Maryland, USA
    \item MPDS (04) – Material Phases Data System, Vitznau, Switzerland – Linus Pauling File (LPF)
    \item ICDD Single Crystal Data (05) – International Centre for Diffraction Data Newtown Square, PA, USA
\end{itemize}
\end{itemize}
